Code to reproduce results in _Semi-parametric generalized linear model for

### 2021-11-19 ManuscriptCode/Manuscript.Rmd
Switch to q-power backbone for misspecified model

  1. Both the beta-binomial and semi-parameteric model are correct - a binomial backbone, which belongs to the exponential family
  2. Only the beta-binomial model is correct - generated from a beta-binomial distribution
  3. Only the semi-parametric model is correct - generated from the spglm, with a zero-inflated beta-binomial backbone
  4. Neither model is correct - generated from a zero-inflated beta-binomial distribution
 * fit the data using our semi-parametric regression model, beta-binomial regression (potentially misspecified distribution), and GEE (marginal only)

@@

# add random response counts from zero-inflated beta-binomial 
add_response_zoibb <- function(cov_data, pstr0, rho){
  # cov_data should come from sim_covariates

  res <- cov_data
  res$NResp <- rzoibetabinom(n = nrow(cov_data),
                             size = cov_data$ClusterSize,
                             prob = cov_data$p_resp,
                             rho = rho,
                             pstr0 = pstr0)
  res$Trt <- "All"
  res$Freq <- 1
  class(res) <- c("CBData", class(res))
  res
}
```

@@

```{r AnalyzeOne}
run_analysis <- function(dat){
  #GEE
  dat_long <- unwrap(dat)
  res_gee <- geeglm(Resp ~ X1 + X2, data=dat_long, id = ID, family=binomial,
                 corstr = "exch")
  #Beta-binomial
  res_bb <- vgam(cbind(NResp, ClusterSize - NResp) ~  X1 + X2, weights = Freq,
                 data=dat, family =  betabinomial(lmu="logitlink"))
  #SPGLM
  op <- options(warn=-1)
  res_spglm <- spglm(cbind(NResp, ClusterSize - NResp) ~ X1 + X2, data=dat,
                   weights  = Freq)
  options(op)

  nms <- names(coef(res_spglm))
  res <- bind_rows(
    GEE = tibble(Term = nms, 
                 Est=coef(res_gee), 
                 SE = sqrt(diag(vcov(res_gee)))),
    BB = tibble(Term = nms,
                Est = coef(res_bb)[-2],      # second value is logit(rho)   
                SE=sqrt(diag(vcov(res_bb)))[-2]),
    SPGLM = tibble(Term = nms,
                   Est = coef(res_spglm), 
                   SE= sqrt(diag(vcov(res_spglm)))),
    .id = "Model")
  res
}
```

@@

```{r ActualSim1, warning=FALSE}
R <- 50

# Scenario 1: binomial PDF
sim1 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0, rho=0);
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")
```

```{r ActualSim2, warning=FALSE}
# Scenario 2: binomial PDF
sim2 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0, rho=0.3);
             run_analysis(simd)}, 
            simplify = FALSE) %>%
   bind_rows(.id="ID")

@@

```{r ActualSim3, warning=FALSE}
# Scenario 3: SPGLM with zero-inflated betabinomial backbone
sim3 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
              add_response_spglm(q0 = dzoibetabinom(0:10, size=10, prob=0.4, rho = 0.3, pstr0 = 0.1));
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

```{r ActualSim4, warning=FALSE}
# Scenario 4: zero-inflated betabinomial 
sim4 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% add_response_zoibb(pstr=0.1, rho=0.3);
             run_analysis(simd)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

@@

# get correct values of parameters on a logit-model scale
correct <- tibble(Term = c("(Intercept)","X1","X2"),
                  Correct = beta_correct)
comb <- left_join(comb, correct, by="Term")

scomb <- comb %>%
  group_by(Generator, Term, Model) %>%
  summarize(bias = mean(Est-Correct, na.rm=TRUE),
            mc.se = sd(Est, na.rm=TRUE),
            se = mean(SE, na.rm=TRUE),
            .groups="drop_last") %>%
  mutate(rel.eff = mc.se / mc.se[Model=="GEE"]) %>%
  ungroup() %>%
  arrange(Generator, Model, Term)
kable(scomb, digits=3)
```

```{r ComparePlot}
ggplot(comb, aes(x=Model, y=Est, colour=Model)) +
  facet_grid(Generator~Term) +
  geom_hline(aes(yintercept=Correct)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(alpha=0.3, height=0)
```

>>>>

  1. Both the beta-binomial and semi-parameteric model are correct - a binomial backbone, which belongs to the exponential family
  2. Only the beta-binomial model is correct - generated from a beta-binomial distribution
  3. Only the semi-parametric model is correct - generated from the spglm, with a q-power backbone
  4. Neither model is correct - generated from a q-power distribution
 * fit the data using our semi-parametric regression model, beta-binomial regression (potentially misspecified distribution), and GEE (marginal only)

@@

# add random response counts from binomial, beta-binomial, or q-power
add_response_param <- function(cov_data, rfun=function(n, cs, p)rbinom(n, size=cs, prob=p)){
  # cov_data should come from sim_covariates

  res <- cov_data
  res$NResp <- rfun(n = nrow(cov_data),
                    cs = cov_data$ClusterSize,
                    p = cov_data$p_resp)
  res$Trt <- "All"
  res$Freq <- 1
  class(res) <- c("CBData", class(res))
  res
}

# random number generation from the q-power pdf
rqpower <- function(n, size, prob, rho){
  size <- rep(size, length.out = n)
  prob <- rep(prob, length.out = n)
  rho <- rep(rho, length.out = n)

  res <- mapply(
    function(s,p,r){
      pdf <- qpower.pdf(p=p, rho=r, n=s)
      sample(0:s, size=1, prob=pdf)},
    size, prob, rho)
  res
}
```

@@

```{r AnalyzeOne}
run_analysis <- function(dat, cs_predict=5){
  #GEE
  dat_long <- unwrap(dat)
  res_gee <- geeglm(Resp ~ X1 + X2, data=dat_long, id = ID, family=binomial,
                 corstr = "exch")
  #Beta-binomial
  res_bb <- vgam(cbind(NResp, ClusterSize - NResp) ~  X1 + X2, weights = Freq,
                 data=dat, family =  betabinomial(lmu="logitlink"))
  #SPGLM
  op <- options(warn=-1)
  res_spglm <- spglm(cbind(NResp, ClusterSize - NResp) ~ X1 + X2, data=dat,
                   weights  = Freq)
  options(op)

  # predict PDF at cluster-size cs_predict & all covariates = 0
  newd <- data.frame(X1 = 0, X2 = 0)
  # beta-binomial
  mns_rho <- predictvglm(res_bb, type="link", newdata=newd, untransform=TRUE)

  fit_bb <- newd %>% 
    crossing(NResp=0:cs_predict) %>%
    mutate(ClusterSize = cs_predict,
           Prob = dbetabinom(NResp, size=cs_predict, prob=mns_rho[1,"mu"], rho=mns_rho[1,"rho"])) %>%
  mutate(Model="BB")

  fit_spglm <- newd %>% 
    crossing(NResp=0:cs_predict) %>%
    mutate(ClusterSize = cs_predict,
           Prob = predict(res_spglm, newdata=., type="prob", 
                          newn=ClusterSize, newevents = NResp),
           Model = "SPGLM")

  nms <- names(coef(res_spglm))
  res <- bind_rows(
    GEE = tibble(Term = nms, 
                 Est=coef(res_gee), 
                 SE = sqrt(diag(vcov(res_gee)))),
    BB = tibble(Term = nms,
                Est = coef(res_bb)[-2],      # second value is logit(rho)   
                SE=sqrt(diag(vcov(res_bb)))[-2]),
    SPGLM = tibble(Term = nms,
                   Est = coef(res_spglm), 
                   SE= sqrt(diag(vcov(res_spglm)))),
    .id = "Model")

  fit <- bind_rows(fit_bb, fit_spglm) %>%
    mutate(Type = "PDF")

  out <- res %>%
    mutate(Type = "Beta") %>%
    bind_rows(fit %>% 
                mutate(Term = as.character(NResp)) %>%
                select(Type, Model, Term, Est=Prob))

  out
}
```

@@

```{r ActualSim1, warning=FALSE}
R <- 100
cs_predict <- 5
p_predict <- 1/(1+exp(-c(beta_correct %*% c(0,0,0))))

# Scenario 1: binomial PDF
sim1 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
                add_response_param(rfun=function(n, cs, p)rbinom(n, size=cs, prob=p));
             run_analysis(simd, cs_predict = cs_predict)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

pdf1 <- tibble(NResp = 0:cs_predict) %>%
  mutate(Correct = dbinom(NResp, size=cs_predict, prob=p_predict))
```

```{r ActualSim2, warning=FALSE}
# Scenario 2: beta-binomial PDF
sim2 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
                add_response_param(rfun=function(n, cs, p)rbetabinom(n, size=cs, prob=p, rho=0.3));
             run_analysis(simd, cs_predict = cs_predict)}, 
            simplify = FALSE) %>%
   bind_rows(.id="ID")

pdf2 <- tibble(NResp = 0:cs_predict) %>%
  mutate(Correct = dbetabinom(NResp, size=cs_predict, prob=p_predict, rho=0.3))
```

```{r ActualSim3, warning=FALSE}
# Scenario 3: SPGLM with q-power backbone
sim3 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
              add_response_spglm(q0 = qpower.pdf(n=10, p=0.4, rho = 0.3));
             run_analysis(simd, cs_predict = cs_predict)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

# simulation-based pdf
pdf3 <- ran.spglm(n=rep(cs_predict, 1e6), means = rep(p_predict, 1e6), 
                  q0 = qpower.pdf(n=10, p=0.4, rho = 0.3)) %>%
  count(NResp) %>%
  mutate(Correct = n/sum(n)) %>%
  select(-n)

@@

```{r ActualSim4, warning=FALSE}
# Scenario 4: q-power 
sim4 <- 
  replicate(R,
            {simd <- sim_covariates(100) %>% 
                add_response_param(rfun=function(n, cs, p)rqpower(n, size=cs, prob=p, rho=0.3));
             run_analysis(simd, cs_predict = cs_predict)}, 
            simplify = FALSE)  %>%
   bind_rows(.id="ID")

pdf4 <- tibble(NResp = 0:cs_predict) %>%
  mutate(Correct = qpower.pdf(n=cs_predict, p=p_predict, rho=0.3))

@@

# get correct values of parameters 
correct_beta <- tibble(Term = c("(Intercept)","X1","X2"),
                  Correct = beta_correct)
correct_pdf <- bind_rows(`Both correct` = pdf1, 
                         `BB correct`= pdf2,
                         `SPGLM correct` = pdf3, 
                         `Neither correct` = pdf4,
                         .id="Generator") %>%
  mutate(Term = as.character(NResp))

comb_beta <- left_join(comb %>% filter(Type == "Beta"), correct_beta, by="Term")

comb_pdf <- left_join(comb %>% filter(Type=="PDF"), 
                      correct_pdf, by=c("Term","Generator"))

scomb <- comb_beta %>%
  group_by(Generator, Term, Model) %>%
  summarize(bias = mean(Est-Correct, na.rm=TRUE),
            mc.se = sd(Est, na.rm=TRUE),
            se = mean(SE, na.rm=TRUE),
            .groups="drop_last") %>%
  mutate(rel.eff = mc.se / mc.se[Model=="GEE"]) %>%
  ungroup() %>%
  arrange(Generator, Model, Term)
kable(scomb, digits=3)
```

```{r ComparePlot}
ggplot(comb_beta, aes(x=Term, y=Est-Correct, colour=Model, shape=Model)) +
  facet_wrap(~Generator) +
  geom_hline(aes(yintercept=0)) +
  stat_summary(fun.data =  mean_sdl, geom="pointrange", position=position_dodge(width=0.4))

```{r PDFplot}
ggplot(comb_pdf, aes(x=NResp, y=Est, colour=Model)) +
  facet_wrap(~Generator) +
  geom_line(aes(y=Correct), col="black") +
  stat_summary(fun.data =  mean_sdl, geom="pointrange", position=position_dodge(width=0.1))
